---
title: "Exercise 5.5.47"
author: "Tony Zhang"
format: html
editor: visual
---

## Exercise 5.2.49:

-   Suppose meiotic drive affects the pollen only and that 80% of the pollen grains from a heterozygote carry the A allele. Ovules are normal and 50% of them carry the A allele. What fraction of offspring from aselfing heterozygote will be heterozygous?

```{r}

numHeteOffSpring = 0
for(n in 1:9999)
for(offSpring in 1:100){
  isA_Pollen_Allele = runif(1) <= 0.8
  isA_Ovule_Allele = runif(1) <= 0.5

  if ((isA_Pollen_Allele && !isA_Ovule_Allele) || (!isA_Pollen_Allele && isA_Ovule_Allele)){
    numHeteOffSpring = numHeteOffSpring + 1
  }
  
}
numHeteOffSpring/9999

```

## Exercise 5.2.50:

-   Suppose meiotic drive affects both pollen and ovules and that 80% of the pollen grains and ovules from aheterozygote carry the A allele. What fraction of offspring from a selfing heterozygote will be heterozygous?

```{r}
numHeteOffSpring = 0
for(n in 1:9999)
for(offSpring in 1:100){
  isA_Pollen_Allele = runif(1) <= 0.8
  isA_Ovule_Allele = runif(1) <= 0.8

  if ((isA_Pollen_Allele && !isA_Ovule_Allele) || (!isA_Pollen_Allele && isA_Ovule_Allele)){
    numHeteOffSpring = numHeteOffSpring + 1
  }
  
}
numHeteOffSpring/9999

```
