---
title: "Type1Type2CellProblem"
author: "Tony Zhang"
format: html
editor: visual
---

## The Problem:

-   For a Type 1 cell, after `t` minutes there is a 1/2 chance that it will remain as a Type 1 cell and there is a 1/2 chance it will turn into a Type 2 cell.

-   For a Type 2 cell, after `t` minutes there is a 3/4 chance it will remain as a Type 2 cell and there is a 1/4 chance it will turn into a Type 1 cell.

-   Problem A: After a long period time, how many Type 1 and Type 2 cells will there be?

-   Problem B: Does the amount of type 1 and 2 cells start out with matter?

## Problem A Solution:

-   let the numerical object of `x_1` denote the amount of type 1 cells and `x_2` denote the amount of type 2 cells.

-   Let `t` denote the amount of time that has passed.

```{r}
x_1 = 6000
x_2 = 6000
#x_1c used so that the function of x_2 uses x_1 for the past iteration of `t` not the current
x_1c = 6000

for (t in 1:9999){
  x_1 = 0.5*x_1 + 0.25*x_2
  x_2 = 0.5*x_1c + 0.75*x_2
  x_1c = x_1
}
```

### Problem A Result:

```{r}
x_1 
x_2 
```

-   It appears that the amount of type 1 cells will be one-third of the total amount of cells and the amount of type 2 cells will be two-thirds of the total amount of cells.

-   Does this hold true for different starting amounts of x_1 and x_2?

## Problem B Solution:

```{r}
x_1 = 0
x_2 = 12000
#x_1c used so that the function of x_2 uses x_1 for the past iteration of `t` not the current
x_1c = 0

for (t in 1:9999){
  x_1 = 0.5*x_1 + 0.25*x_2
  x_2 = 0.5*x_1c + 0.75*x_2
  x_1c = x_1
}
```

## Problem B Result:

```{r}
x_1 
x_2
```

-   It resulted in the same answer, so I conclude that it does not matter.
