---
title: "graphs"
author: "Montana Ferita"
date: "2024-02-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading data

This package is required to be able to read the data

```{r}
library(readr)
library(ggplot2)
```

```{r}
data <-read.csv("C:/Users/Zhang/Documents/GitHub/SRICode_Repo/RGraphsInto/104lineage.csv")
```

if you want to view part of the data use the head function

```{r}
head(data)
```

if you want to remove some columns use the line below

```{r}
data <- data[ -c(6:10,19,21) ]
head(data)
```

# some basic plots

You can use the plot function which you say the data set you are using and then the columns you want to look at separated by a dollar sign

```{r}
plot(data$Frame,data$Tracking_)
```

Another way you can create graphs is to use ggplot

```{r}
# Create a color palette with as many colors as unique Tracking_IDs
color_palette <- rainbow(length(unique(data$Tracking_ID)))

# Create a ggplot with geom_point
ggplot(data = data, aes(x = Dry_Mass, y = mCherry_Mean_Intensity, color = factor(Tracking_ID))) +
  geom_point() +
  scale_color_manual(values = color_palette) +
  labs(title = "Scatter Plot with Distinct Colors for Tracking_ID")


```

### Bar Graph of Intensities for different cells

```{r}
boxplotV(mCherry_Integrated_Intensity~Tracking_ID, data)
```
