---
title: "Crowding_Intensity_Simulation"
author: "Tony"
date: "2024-07-09"
output: pdf_document
---

```{r setup, include=FALSE}
library(MASS)
library(ggplot2)
library(tidyverse)
```

# Reading in Data

```{r}
raw_df = read.csv('Data/ken6hrs.csv')


# Setting up starting conditions of the simulation
raw_df$Normalized_Intensity = raw_df$Integrated.Intensity / raw_df$Dry.Mass

raw_df <- raw_df %>%
  group_by(Tracking.ID) %>%
  mutate(Normalized_Intensity_Change = lead(Normalized_Intensity) / Normalized_Intensity)

sim_df = subset(raw_df, select = c(Frame, Normalized_Intensity, Position.X, Position.Y), subset = Frame == 1)

row.names(sim_df) <- NULL

```

# Simulation

## Cell Count

-   This simulation will attempt to stay consistent with the number of cells in the actual experiment.
    -   However, the number of cells which undergo immigration, emigration, death and division each Frame is unknown.
    -   This experiment will ignore all but division so will fit an exponential curve to cell counts to ensure cell count never decreases.

```{r}
cell_counts_df = as.data.frame(table(raw_df$Frame))

names(cell_counts_df) = c("Frame", "Cell_Count")

cell_counts_df$Frame <- as.numeric(as.character(cell_counts_df$Frame))

ggplot(cell_counts_df, aes(x = Frame, y = Cell_Count)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(min(cell_counts_df$Frame), max(cell_counts_df$Frame), by = 10)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Appears to be a steep drop off in cells at Frame 261, so will only simulate to Frame 261.

```{r}
cell_counts_df = cell_counts_df[cell_counts_df$Frame <= 261, ]

# Fitting an exponential model to the count
model <- lm(log(Cell_Count) ~ Frame, data = cell_counts_df)

#ceiling rounds up, as cannot have 0.24 of a cell
cell_counts_df$predicted = ceiling(exp(predict(model, cell_counts_df)))



ggplot(cell_counts_df, aes(x = Frame, y = Cell_Count)) + 
  geom_point() +
  scale_x_continuous(breaks = seq(min(cell_counts_df$Frame), max(cell_counts_df$Frame), by = 10)) +
  geom_line(data = cell_counts_df, aes(x = Frame, y = predicted, ), color = 'blue', size = 1) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  labs(title = "Exponential Fit to Cell Counts")
```

```{r}
#minimum amount of time required to divide
maturation_time = 6.0

#will assume all starting cells are capable of division
sim_df$duration_hr = maturation_time

cell_counts_df <- cell_counts_df %>%
  mutate(count_increase = lead(predicted) - predicted)
```

## Intensity:

```{r}
clean_df = na.omit(raw_df)

mean_change = mean(clean_df$Normalized_Intensity_Change)
sd_change = sd(clean_df$Normalized_Intensity_Change)

ggplot(clean_df, aes(x = log(Normalized_Intensity_Change)))+
  geom_density() +
  stat_function(fun=dnorm, args=list(mean=mean_change, sd=sd_change), color="red", size=1) +

  labs(title = "Log Distribution of Change in Intensities")
```

## Results

```{r}
for(i in 1:260){
  curr_df = sim_df[sim_df$Frame == i, ]
  curr_df$Frame = i+1
  
  cell_increase = cell_counts_df$count_increase[i]
  
  if(cell_increase > 0){
    #Randomly choose cells to divide
    filtered_df = curr_df[curr_df$duration_hr >= maturation_time, ] #cells must be mature enough
  
    chosen_indices = sample(nrow(filtered_df), cell_increase) #randomly choose indices of cells to divide
    chosen_cells = filtered_df[chosen_indices, ]
    
    original_sampled_indices = which(curr_df$duration_hr >= maturation_time)[chosen_indices] #find the indices of the chosen cells to divide
    curr_df = curr_df[-original_sampled_indices, ] #remove these cells from the df
  
    for(i in 1:nrow(chosen_cells)){
      left_child = data.frame(
                            Frame = chosen_cells[i, "Frame"], 
                            Normalized_Intensity = chosen_cells[i, "Normalized_Intensity"],
                            Position.X = chosen_cells[i, "Position.X"] - 20,
                            Position.Y = chosen_cells[i, "Position.Y"],
                            duration_hr = 0
                            )
      right_child = data.frame(
                            Frame = chosen_cells[i, "Frame"], 
                            Normalized_Intensity = chosen_cells[i, "Normalized_Intensity"],
                            Position.X = chosen_cells[i, "Position.X"] + 20,
                            Position.Y = chosen_cells[i, "Position.Y"],
                            duration_hr = 0
                            )
    
      children = rbind(left_child, right_child)
    
      curr_df = rbind(curr_df, children) #add children to current frame
      }
    
    }
  
  
  curr_df$Normalized_Intensity = curr_df$Normalized_Intensity * exp(rnorm(n = nrow(curr_df), mean_change, sd_change))
  curr_df$duration_hr = curr_df$duration_hr + 0.25
  
  
  
  
  
  sim_df = rbind(sim_df, curr_df)#add the new frame to the simulated data
  
}

row.names(sim_df) <- NULL
```

```{r}
write.csv(sim_df, file = "simulation01.csv", row.names = FALSE)
```
